Placeholder Tex Doc

### 2019-01-06 tex_files/placeholder.tex
...

\title{Placeholder Tex Doc}
\author{Conor Dooley}
\maketitle
\chapter{Background Review}
\section{Brief Overview}
\doublespacing
In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the forefront. %TODO cite?
The main drivers of performance in Systems-On-Chip are the number of transistors on a chip, the clocking frequency and parallelism. However the latter will not be addressed in this assessment, partly as the performance of the paralellised operation is largely governed by the number of transistors and clock frequency to begin with. %TODO Excessive?
As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last 10 years \cite{ross2008cpu}.

@@

\section{The Impact of Clocking Errors}

In Figure \ref{fig:eldar_why_precise_clocking} the data path between two synchronously clocked registers is shown, with the circuit's function being carried out by the combinatorial network between the registers.
Each register has a setup time, which represents the amount of time that the input value to a register must remain constant before the clock edge, and a hold time, the time for which the input must remain constant after a clock edge.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}
	\label{fig:eldar_why_precise_clocking}
\end{figure}\newline
A lack of synchronisation between the clock edges will manifest itself as a time difference between the clocking events at both registers, $\Delta T = t^i_k - t^f_k$. $\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process that is normally modelled as a Gaussian random variable. If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby having the same effect as a reduction in clocking frequency.
The most common sources of clock error are caused by mismatches which usually stem from production such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network. These all manifest themselves as skew, while the noise in active components will appear as jitter in the clock signal.

\section{Traditional Solutions}

A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques. The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees. %TODO insert figure
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
	\label{fig:trees}
\end{figure}
While on the surface these appear simple the task of obtaining an exact matching is, in practise, the limiting factor in this design. Even if the clock distribution system is geometrically symmetrical production mismatches in either active or passive components will lead to a skew that varies from part to part. In order to minimise the impact of production tolerances the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible. However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.

@@

\section{ADPLL Networks}

As a PLL network is an analog circuit its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}. An alternative design that is more suitable for current fabrication techniques eschews from using analog components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital PLL. A 4x4 ADPLL network was designed and prototyped in 65 nm CMOS by Zianbetov and Shan in order to test the suitability of the technique as a clock distributor \cite{zianbetov2013phd,shan2014phd}. %TODO fix phd references
In this design the oscillators are coupled using digital phase comparators which attempt to measure the phase error between two oscillators with each non edge node being connected to four neighbours. Figure \ref{fig:eldar_node} shows high level detail of the architecture of both the entire clocking system and that of an individual node in the design. A digital PLL network has the additional advantages over its analog counterparts that it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable and has a significantly greater immunity to perturbations inherent to its digital nature \cite{zianbetov2013phd}. This last advantage is of particular use in a digital environment, as otherwise there is potential for clock degradation resulting from switching of transistors.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../ccirc_2013_arch}
	\caption{Architecture of the ADPLL network and of a single node \cite{zianbetov2013distributed}.}
	\label{fig:eldar_node}
\end{figure}
Looking at the design of a given node it is notable that the function carried out by the Error Combiner  is akin to a average therefore as mentioned by Pratt and Nyugen there is potential to settle in a stable ``mode'' in which the oscillators are not synchronised. In their paper they present a method where initial start-up is performed uni-directionally and once all nodes are close full connectivity can be restored \cite{pratt1995distributed}. In creating an entirely digital system Zianbetov and Shan could exploit the reconfigurability and and implement this system at start-up.

@@

@@

>>>>

\title{Placeholder Tex Doc}
\author{Conor Dooley}
\maketitle
\doublespacing
\chapter{Background Review}
\section{Brief Overview}
In a world where the demand for high performance hand-held computing devices continues to grow and the prevalence of ``smart'' devices is increasing the problem of maintaining the steady gain in performance of Systems-On-Chip remains at the forefront. %TODO cite?
The main drivers of performance in Systems-On-Chip are the number of transistors on a chip, the clocking frequency and parallelism. However the latter will not be addressed in this assessment, partly as the performance of the paralellised operation is largely governed by the number of transistors and clock frequency to begin with. %TODO Excessive?
As the number of transistors on a chip has increased roughly following Moore's Law the increase in clock frequency has not been able to follow a similar linear trajectory, which clock frequency having remained roughly equivalent for the last number of years \cite{ross2008cpu}.

@@

\section{The Impact of Clocking Errors}
In Figure \ref{fig:eldar_why_precise_clocking} the data path between two synchronously clocked registers is shown, with the circuit's function being carried out by the combinatorial network between the registers.
Each register has a setup time, which represents the amount of time that the input value to a register must remain constant before the clock edge, and a hold time, the time for which the input must remain constant after a clock edge.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013phd}.}
	\label{fig:eldar_why_precise_clocking}
\end{figure}\newline
A lack of synchronisation between the clock edges will manifest itself as a time difference between the clocking events at both registers, $\Delta T = t^i_k - t^f_k$. $\Delta T$ is considered to be ergodic and can be described by an average deviation called skew and random process that is normally modelled as a Gaussian random variable. If $\Delta T$ is negative this reduces the time available for the intervening combinatorial network thereby having the same effect as a reduction in clocking frequency.
The most common sources of clock error are caused by mismatches which usually stem from production such as differences in the length of clocking paths, buffer delays or in the parameters of either active or passive components in the clock distribution network. These all manifest themselves as skew, while the noise in active components will appear as jitter in the clock signal.

\section{Traditional Solutions}
A number of traditional solutions exist which provide GSLS clocking systems, using a variety of techniques. The most simple of these use the clock distribution network's symmetry in order to deliver the clock in phase at all locations in the chip and are named according to their geometry with the main variants being branch-trees, X-trees or H-trees. %TODO insert figure
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../eldar_why_precise_clocking}
	\caption{Data Flow in a Clocked System \cite{zianbetov2013distributed}.}%TODO insert figure
	\label{fig:trees}
\end{figure}
While on the surface these appear simple the task of obtaining an exact matching is, in practise, the limiting factor in this design. Even if the clock distribution system is geometrically symmetrical production mismatches in either active or passive components will lead to a skew that varies from part to part. In order to minimise the impact of production tolerances the dimensions of components in the distribution network can be increased, thus reducing the relative variation possible. However this has the impact of increasing the power consumption of the distribution network \cite{tiwari1998reducing}.
Designs do exist in which the electrical lines used in the tree network are replaced by waveguides for optical signals, which negates the electrical losses however this technique suffers from being a newer technology and as such is expensive to realise \cite{haurylau2006chip}.

@@

\section{ADPLL Networks}
As a PLL network is an analog circuit its integration in a modern IC is a barrier to usage and it has not been used in any commercial designs \cite{zianbetov2013distributed}. An alternative design that is more suitable for current fabrication techniques eschews from using analog components and instead implements the network of controlled oscillators using only digital circuitry, hence the name All-Digital PLL. A 4x4 ADPLL network was designed and prototyped in 65 nm CMOS by Zianbetov and Shan in order to test the suitability of the technique as a clock distributor \cite{zianbetov2013phd,shan2014phd}. %TODO fix phd references
In this design the oscillators are coupled using digital phase comparators which attempt to measure the phase error between two oscillators with each non edge node being connected to four neighbours. Figure \ref{fig:eldar_node} shows high level detail of the architecture of both the entire clocking system and that of an individual node in the design. A digital PLL network has the additional advantages over its analog counterparts that it can benefit from advancements in digital circuit design suites, be reconfigurable and programmable and has a significantly greater immunity to perturbations inherent to its digital nature \cite{zianbetov2013phd}. This last advantage is of particular use in a digital environment, as otherwise there is potential for clock degradation resulting from switching of transistors.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../ccirc_2013_arch}
	\caption{Architecture of the ADPLL network and of a single node \cite{zianbetov2013distributed}.}
	\label{fig:eldar_node}
\end{figure}
Looking at the design of a given node it is notable that the function carried out by the Error Combiner  is akin to a average therefore as mentioned by Pratt and Nyugen there is potential to settle in a stable ``mode'' in which the oscillators are not synchronised. In their paper they present a method where initial start-up is performed uni-directionally and once all nodes are close full connectivity can be restored \cite{pratt1995distributed}. In creating an entirely digital system Zianbetov and Shan could exploit the reconfigurability and and implement this system at start-up.

@@

\section{My Contribution}
The planned contribution of this project is to create an FPGA based prototyping platform for the testing of network parameters and architectural variations. An FPGA as a prototyping tool for networks was previously utilised by Zianbetov and Shan as the final stage before the creation of a custom chip as it allowed for ``a partial validation of the ASIC design, through a verification of the essential functional features of the designed clocking network \cite{zianbetov2013phd,shan2014phd}. They however note that there are some fundamental differences between the two environments in particular that the FPGA based prototype is unable to create the mixed signal circuits used in an ASIC resulting in alternative designs being mandated for certain components such as the Time-to-Digital Converter. Similarly most commericially available FPGA boards were noted to have clock frequencies well below the gigahertz levels possible in ASIC designs. In spite of these problems the FPGA as a prototyping solution is valuable as it allows for hardware based testing of architectures and the simulation of non trivial circuit behaviours.

The novelty of this particular implementation is that in the case of Zianbetov and Shan only a small network was simulated on an FPGA and at a frequency close to 50 kHz, however this project will attempt to significantly increase the frequency of operation to the MHz range as well as expanding the size of the network, from 16 oscillators to potentially 100. This is done with the aim of assisting work being undertaken by a number of researchers in UCD who are working on aspects of ADPLL networks who would stand to benefit from such a platform.

@@

The first stage of this project was the creation of an ADPLL design that would be used for the network. Due to limitations of the connections available on the FPGA evaluation board used to this point, a Digilent/Xilinx Nexys 4, signals at frequencies significantly beyond 5 MHz suffered from major degradation as the connections were not designed for such frequencies. Thus a target frequency was set at 5 MHz while this board was in use. After initial research into potential oscillator designs and what would be feasible given the constraints of the particular platform design began with the aforementioned oscillator.
\section{Controllable Oscillator}
In my research into potential designs for numerically controllable oscillators that were viable on an FPGA I came across two main options. The first of these was the ring oscillator, constructed from an odd number of inverters strung together in a chain \cite{predraig}. In this case the frequency of operation is determined by the number of oscillators in the chain as the delay through the individual inverters determines the half period of the signal. This period can then be modified by adding or removing inverters in a manner that maintains the odd number of total inverters. The propagation time through an individual inverter is dependant on the layout of the ring and thus this design must be used with some care.\\
The second option is one that I was aware of in advance this project's beginning. A phase accumulator makes use of the overflow property of a counter, and has a frequency determined by the width of the counter and the frequency at which it is clocked. The oscillator operates as follows: Once per clock cycle the counter will increment by a number $k$ until overflow is reached at $2^n-1$ where $n$ is the bit width of the counter. By controlling the magnitude of $k$ added on each cycle of the driving clock the frequency of this oscillator can be adjusted. The most significant bit of this counter can form the clock signal as it has an approximately 50\% duty cycle.\\
As it is not reliant on the clock of FPGA the ring oscillator design is a more interesting design in the context of simulation, but the phase accumulator is significantly easier to manage due to being more easily simulatable. The decision was made to create both designs in order to keep options open.\\
The only other option I managed to come across for an oscillator on an FPGA was using Xilinx proprietary \texttt{IODELAY} blocks, which have a permanently programmable delay at runtime \ref{IODELAY}. This could be coupled with an odd number of inverters in order to implement a quasi ring oscillator. However as these are IO blocks their usage is not particularly conducive to the creation of a network of ADPLLs.\\
The phase accumulator was designed to have a bit width of 12, all of which could be set as the control code. As the maximum frequency of operation that had no timing violations was 258 MHz the frequency step of this oscillator design was 62.988 kHz and before integration into an ADPLL had a frequency range from 62.988 kHz to just under half the clock frequency at 128.94 MHz. Figure \ref{fig:my_pa} details the simple construction of this oscillator.\\
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{../phase_accum}
	\caption{Phase Accumulator Block Diagram.}
	\label{fig:my_pa}
\end{figure}
As originally implemented for testing purposes the inverter ring oscillator had a propagation time per inverter of 315 picoseconds, thus each inverter would add 0.63 nano seconds to the total period, and had 4 control bits. In order to obtain a 5 MHz output from this type of ring oscillator a ring of of 317 inverters is required, and to set this as the centre point of the frequency range the total number of inverters was 325. At 5 MHz the frequency step of this design was calculated as:
\begin{align*}
T_{nom} = 200\times10^{-9}\text{ ns ,}T_{next} = 201.26\times10^{-9}\text{ ns} \\
F_{nom} = 5\times10^{6}\text{ Hz ,}F_{next} = 4.968697\times10^{-9}\text{ Hz}
\end{align*} 
\begin{equation*}
\Delta F_{RO} = 31.302\text{ kHz}
\end{equation*}
%TODO look into more detail on yoke.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{../ring_osc}
	\caption{Simplified Ring Oscillator Block Diagram.}
	\label{fig:my_ring_osc}
\end{figure}
\chapter{Future Work}
